Percussive Synthesizer Generation with Virtual Listeners

### 2020-09-30 templates_aimusic2020/paper.tex
Updates from Overleaf
some writing errors corrected as pointed out by reviewers

\documentclass[runningheads,a4paper]{llncs}
\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{url}
\usepackage{apacite}
%added 
\usepackage[inline]{enumitem}    
\usepackage{subcaption}
\usepackage{caption}
\usepackage{threeparttable} %annotations for table
\usepackage{csquotes}
\usepackage{appendix}
\usepackage{comment}
\usepackage{multirow}
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

@@

\section{Related Works}
Numerous deep neural network models have been proposed for the purpose of signal generation in recent years~\cite{nsynth2017,yamamoto2020parallel,oord2017parallel,yee2018automatic,ramires2020neural}. WaveGans and WaveNet have been subject to significant improvements and experiments since their proposal~\cite{nsynth2017,yamamoto2020parallel,oord2017parallel}. Particularly relevant works are the utilization of variational autoencoders (VAE's) for generation of percussive samples~\cite{aouameur2019neural} and generation of percussive sounds by decoding a small set of latent features~\cite{ramires2020neural}. Automatic programming~\footnote{unsupervised generation of source code or programs towards a goal} of virtual synthesizers has long been a topic of interest. In early 2000s, Interactive Genetic Algorithms (IGA's) were utilized for the generation of new sounds with various sound-engines~\cite{johnson1999exploring,dahlstedt2001creating}. A more recent work by Yee-King et al.~\cite{yee2018automatic} used Long short-Term Memory (LSTM) models and heuristic search to find the exact parameters used to create a group of sounds. The sounds approximated were made by the same virtual synthesizer and not with an external source. Esling et al. used  over 10,000 a VST synthesizer's presets to learn a parameter space which can be sampled for creation of new audio~\cite{esling2019universal}. Here, we work towards the approximation of percussive sounds with no prior knowledge about the sonic capabilities of a synthesizer.

\section{Virtual Synthesizer Design}
\begin{table}[tbp]
\centering
\resizebox{\columnwidth}{!}{\begin{tabular}{ |p{4cm}|p{4cm}|p{4cm}| } 
\hline
Parameters & Value Range & notes and constraints\\
\hline \hline
Attack & 0-3 & A-D-S-R values relative\\
Decay & 0-3 & relative to A-S-R\\
Sustain & 0-3 & relative to A-D-R\\
Release & 0-3 & relative to A-D-S\\
OSC type & sine,square,saw & -\\
IsNoise & boolean & generate noise using \newline cloud of waveform\\
Length & 0-1 second & - \\
StartTime & 0-1 second & Length+Start$<$1\\
Amplitude & 0.1-1 & 1 = max amplitude\\
Pitches(notes) & list of pitches &  range of C0(16.35hz) to B9 \\
HP filter Cuttoff & 0-20000hz & -\\
LP filter Cuttoff & 20000-HP & never lower than HP cutoff\\
Filter Order & 4,8,16 & butterworth filter order \\
\hline
\end{tabular}}
\caption{Synthesizer submodule parameters. There are $10^{15}$ unique programs are possible. Each synthesizer can contain any number of submodules}
\label{table:submodule_params}
\end{table}
To create sounds, we need digital synthesizers. We use classical DSP to build our synthesizer, which allows for quick, offline, and parallel generation of audio signals without the usage of GPUs. We made extensive use of Pippi\footnote{https://github.com/luvsound/pippi} and SciPy~\cite{jones2001scipy} libraries. Our virtual synthesizer contains a set of one or more submodules. Each submodule is a self-contained noise making unit and creates signals by taking the steps depicted in figure~\ref{fig:submodule}. Submodules have identical sets of parameters, but widely different outputs can be achieved depending on the values assigned. The set of required parameters for each submodule is highlighted in Table~\ref{table:submodule_params}. The sonic output of the virtual synthesizer is the normalized addition of the output of its submodules. Our implementation of a synthesizer can have any number of submodules. We call the number of submodules in each virtual synthesizer the \textit{stack size}. We call the sets of parameter values that characterize a synthesizer's submodules a \textit{program} (analogous to presets and submodules for a VST).  
 \begin{figure}[tbp]
    \begin{center}
    % \textbf{Synthesizer SubModule }
    \makebox[\textwidth]{
    \fbox{\includegraphics[width=1\linewidth]{templates_aimusic2020/images/synthesizer_block.pdf}}}
    \end{center}
    \caption{High level representation of pre-rendering steps for each submodule. Each Synthesizer contains 1 or more these submodules. Synthesizer programs set the number of these submodules and their parameters. The output of a synthesizer is the normalized addition of all its submodule outputs. 
    }
\label{fig:submodule}
\end{figure}

@@

\caption{Fleiss’ kappa coefficients as a measurement of agreement between persons (H+H) and persons with models.  }
\end{table}
High accuracy in detection of drums vs not-drums does not necessarily make an ear more suitable for our generative system. Perfect detection of drums vs not-drums would create a system which generates nothing.  We cannot know the true size of $\mathcal{H^{+}}$ (see Figure~\ref{fig:ven_data}) without exhaustive manual hearing tests. We conduct manual hearing tests of two generative systems. For \decfirst, a network which learns from spectrogram data utilizing CNN and LSTM layers is used. For \decsecond, we categorize the drum types with 3 different models: FC (Fully-Connected architecture and spectrograms for learning), CNN (CNN and LSTM layers with spectrograms), and E/F (Fully-Connected architecture with frequency and envelope features). The MEM system uses our best MEM, which simultaneously classifies sounds as drums and categorizes them.

@@

%==============
\chapter{Survey Details}
\label{appendix:surveys}
\begin{table}[h!]
  \begin{tabular}{|c | c | c | c | c | c | c|} 
  \hline
  Drop Rule & Size & H+H & H+FC & H+CNN & H+E/F & 3 models \\ [0.5ex] 
  \hline
  No Drops & 257 &0.37 & 0.35 & 0.36 & 0.36 & 0.28\\ 
  \hline
  Assigned \enquote{Bad} By Both & 236 & 0.31 & 0.37 & 0.37 & 0.38 & 0.30 \\
  \hline
  Assigned \enquote{Bad} By Either & 180 & 0.47 & 0.50 & 0.48 & 0.48 &  0.34 \\
  \hline
  Assigned \enquote{Bad} or \enquote{Other} By Either & 154 & 0.47 & 0.59 & 0.54 & 0.50 &  0.35 \\
  \hline
 \end{tabular}
 \caption{\label{kappa_table_TPE}Table of Fleiss' kappa coefficient to measure the degree of agreement between persons (H+H), persons with FC model (H+FC), persons with CNNLSTM model, persons with all models (H+E/F), and between the 3 models. } % AH: Which decision is this?
 \end{table}
\begin{table}[h!]
 \begin{tabular}{|c| p{2cm} | p{2cm}| p{2cm}|} 
 \hline
 Drop Rule & Size & H+H & H+MEM \\
 \hline
 No Drop & 300 & 0.34 & 0.25\\ 
 \hline
 Assigned Bad By Both & 249 & 0.20 & 0.26 \\
 \hline
 Assigned Bad By Either & 151 & 0.46 &  0.47 \\
 \hline
Assigned \enquote{Bad} or \enquote{Other} By Either  & 120 & 0.62   &  0.59 \\
 \hline
\end{tabular}
\caption{\label{kappa_table_MEM}Table of Fleiss' kappa coefficient to measure the degree of agreement between persons (H+H) and persons with MEM (H+MEM).}
\end{table}
% \subsubsection{TPE Survey of 257 Samples:}
% \begin{itemize}
%     \item With 6 categorization groups, responders had the same categorization in 47\% of cases.
%     \item The agreement between the responders and AVG was 44\% and 47\%.
%     \item Of 257 samples, the responders agreed with FC, CNNLSTM and E/F respectively in 78, 76, and 46 of cases. The loss of spectrum data is likely the cause of E/F model's noticeably worse performance. 
%     \item The survey brings into question the reliability of our phase 1 model, as 30\% of the generated samples were deemed not percussive by at least 1 reviewer and 8\% by both reviewers
%     \item The task of categorizing synthetic drums is difficult. Survey shows that the scale of agreement within persons as well as between persons and various model combinations is moderate at best, even after removal of \enquote{bad} samples.  While the same models can easily achieve 98+ percent accuracy when tested on sounds from physical drum samples. 
%     \item While there is much room for improvement, our pipelines can generate and categorize drums and percussive sounds with a promising degree of success. 
% \end{itemize}

>>>>

\documentclass[runningheads,a4paper]{llncs}
\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{url}
\usepackage{apacite}
%added 
\usepackage[inline]{enumitem}    
\usepackage{subcaption}
\usepackage{caption}
\usepackage{threeparttable} %annotations for table
\usepackage{csquotes}
\usepackage{appendix}
\usepackage{comment}
\usepackage{multirow}
\usepackage[T1]{fontenc}
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

@@

\section{Related Works}
Numerous deep neural network models have been proposed for the purpose of signal generation in recent years~\cite{nsynth2017,yamamoto2020parallel,oord2017parallel,yee2018automatic,ramires2020neural}. WaveGans and WaveNet have been subject to significant improvements and experiments since their proposal~\cite{nsynth2017,yamamoto2020parallel,oord2017parallel}. Particularly relevant works are the utilization of variational autoencoders (VAE's) for generation of percussive samples~\cite{aouameur2019neural} and generation of percussive sounds by decoding a small set of latent features~\cite{ramires2020neural}. Automatic programming~\footnote{unsupervised generation of source code or programs towards a goal} of virtual synthesizers has long been a topic of interest. In early 2000s, Interactive Genetic Algorithms (IGA's) were utilized for the generation of new sounds with various sound-engines~\cite{johnson1999exploring,dahlstedt2001creating}. A more recent work by Yee-King et al.~\cite{yee2018automatic} used Long short-Term Memory (LSTM) models and heuristic search to find the exact parameters used to create a group of sounds. The sounds approximated were made by the same virtual synthesizer and not with an external source. Esling et al. used  over 10,000 VST synthesizer presets to learn a parameter space which can be sampled for creation of new audio~\cite{esling2019universal}. Here, we work towards the approximation of percussive sounds with no prior knowledge about the sonic capabilities of a synthesizer.

\section{Virtual Synthesizer Design}
\begin{table}[tbp]
\centering
\resizebox{\columnwidth}{!}{\begin{tabular}{ |p{4cm}|p{4cm}|p{4cm}| } 
\hline
Parameters & Value Range & notes and constraints\\
\hline \hline
Attack & 0-3 & A-D-S-R values relative\\
Decay & 0-3 & relative to A-S-R\\
Sustain & 0-3 & relative to A-D-R\\
Release & 0-3 & relative to A-D-S\\
OSC type & sine,square,saw & -\\
IsNoise & boolean & generate noise using \newline cloud of waveform\\
Length & 0-1 second & - \\
StartTime & 0-1 second & Length+Start$<$1\\
Amplitude & 0.1-1 & 1 = max amplitude\\
Pitches(notes) & list of pitches &  range of C0(16.35hz) to B9 \\
HP filter Cutoff & 0-20000hz & -\\
LP filter Cutoff & 20000-HP & never lower than HP cutoff\\
Filter Order & 4,8,16 & butterworth filter order \\
\hline
\end{tabular}}
\caption{Synthesizer submodule parameters. $10^{15}$ unique programs are possible. Each synthesizer can contain any number of submodules}
\label{table:submodule_params}
\end{table}
To create sounds, we build digital synthesizers. We use classical DSP to build our synthesizer, which allows for quick, offline, and parallel generation of audio signals without the usage of GPUs. We made extensive use of Pippi\footnote{https://github.com/luvsound/pippi} and SciPy~\cite{jones2001scipy} libraries. Our virtual synthesizer contains a set of one or more submodules. Each submodule is a self-contained noise making unit and creates signals by taking the steps depicted in figure~\ref{fig:submodule}. Submodules have identical sets of parameters, but widely different outputs can be achieved depending on the values assigned. The set of required parameters for each submodule is highlighted in Table~\ref{table:submodule_params}. The sonic output of the virtual synthesizer is the normalized addition of the output of its submodules. Our implementation of a synthesizer can have any number of submodules. We call the number of submodules in each virtual synthesizer the \textit{stack size}. We call the sets of parameter values that characterize a synthesizer's submodules a \textit{program} (analogous to presets and submodules for a VST).  
 \begin{figure}[tbp]
    \begin{center}
    % \textbf{Synthesizer SubModule }
    \makebox[\textwidth]{
    \fbox{\includegraphics[width=1\linewidth]{templates_aimusic2020/images/synthesizer_block.pdf}}}
    \end{center}
    \caption{High level representation of pre-rendering steps for each submodule. Each Synthesizer contains 1 or more these submodules. Synthesizer programs set the number of these submodules and their parameters. The output of a synthesizer is the normalized addition of all its submodule outputs. 
    }
\label{fig:submodule}
\end{figure}

@@

\caption{Fleiss’ kappa coefficients as a measurement of agreement between persons (H+H) and persons with models.  }
\end{table}
High accuracy in detection of organic drums vs synthetic sounds does not necessarily make an ear more suitable for our generative system. Perfect detection of drums vs not-drums would create a system which generates nothing.  We cannot know the true size of $\mathcal{H^{+}}$ (see Figure~\ref{fig:ven_data}) without exhaustive manual hearing tests. We conduct manual hearing tests of two generative systems. For \decfirst, a network which learns from spectrogram data utilizing CNN and LSTM layers is used. For \decsecond, we categorize the drum types with 3 different models: FC (Fully-Connected architecture and spectrograms for learning), CNN (CNN and LSTM layers with spectrograms), and E/F (Fully-Connected architecture with frequency and envelope features). The MEM system uses our best MEM, which simultaneously classifies sounds as drums and categorizes them.

@@

%==============
\chapter{Survey Details}
\label{appendix:surveys}
\begin{table}[h!]
  \begin{tabular}{|c | c | c | c | c | c | c|} 
  \hline
  Drop Rule & Size & H+H & H+FC & H+CNN & H+E/F & 3 models \\ [0.5ex] 
  \hline
  No Drops & 257 &0.37 & 0.35 & 0.36 & 0.36 & 0.28\\ 
  \hline
  Assigned \enquote{Bad} By Both & 236 & 0.31 & 0.37 & 0.37 & 0.38 & 0.30 \\
  \hline
  Assigned \enquote{Bad} By Either & 180 & 0.47 & 0.50 & 0.48 & 0.48 &  0.34 \\
  \hline
  Assigned \enquote{Bad} or \enquote{Other} By Either & 154 & 0.47 & 0.59 & 0.54 & 0.50 &  0.35 \\
  \hline
 \end{tabular}
 \caption{\label{kappa_table_TPE}Table of Fleiss' kappa coefficient to measure the degree of agreement between persons (H+H), persons with FC model (H+FC), persons with CNNLSTM model, persons with all models (H+E/F), and between the 3 models. } % AH: Which decision is this?
 \end{table}
\begin{table}[h!]
 \begin{tabular}{|c| p{2cm} | p{2cm}| p{2cm}|} 
 \hline
 Drop Rule & Size & H+H & H+MEM \\
 \hline
 No Drop & 300 & 0.34 & 0.25\\ 
 \hline
 Assigned Bad By Both & 249 & 0.20 & 0.26 \\
 \hline
 Assigned Bad By Either & 151 & 0.46 &  0.47 \\
 \hline
Assigned \enquote{Bad} or \enquote{Other} By Either  & 120 & 0.62   &  0.59 \\
 \hline
\end{tabular}
\caption{\label{kappa_table_MEM}Table of Fleiss' kappa coefficient to measure the degree of agreement between persons (H+H) and persons with MEM (H+MEM).}
\end{table}
